%%  ************    LibreSilicon's 1st TestWafer    *******************
%%
%%  Organisation:   Chipforge
%%                  Germany / European Union
%%
%%  Profile:        Chipforge focus on fine System-on-Chip Cores in
%%                  Verilog HDL Code which are easy understandable and
%%                  adjustable. For further information see
%%                          www.chipforge.org
%%                  there are projects from small cores up to PCBs, too.
%%
%%  File:           PearlRiver/Documents/LaTeX/pearlriver.tex
%%
%%  Purpose:        Top Level File for 1st Test Wafer Documentation
%%
%%  ************    LuaLaTeX with circdia.sty package   ***************
%%
%%  ///////////////////////////////////////////////////////////////////
%%
%%  Copyright (c) 2018 by chipforge <hsank@nospam.chipforge.org>
%%  All rights reserved.
%%
%%      This Standard Cell Library is licensed under the Libre Silicon
%%      public license; you can redistribute it and/or modify it under
%%      the terms of the Libre Silicon public license as published by
%%      the Libre Silicon alliance, either version 1 of the License, or
%%      (at your option) any later version.
%%
%%      This design is distributed in the hope that it will be useful,
%%      but WITHOUT ANY WARRANTY; without even the implied warranty of
%%      MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
%%      See the Libre Silicon Public License for more details.
%%
%%  ///////////////////////////////////////////////////////////////////
\documentclass[a4paper,english,twoside]{article}
\usepackage{fontspec}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{babel}
\usepackage{luatexja-fontspec}
\usepackage[digital,srcmeas,semicon]{circdia}
\usepackage[left=2cm,right=2cm,top=1.5cm,bottom=3cm]{geometry}

\defaultfontfeatures{Ligatures=TeX}
%\setmainfont{Linux Libertine 0}
\setmainjfont{FandolSong}
%\setsansfont{Kurier}
%\setsansfont{Menlo}

\title{PearlRiver -- LibreSilicon's 1st Test Wafer (珠江芯片一号)}
\author{Hagen Sankowski}
\date{\today}

\makeindex  % usefull for ToC
\setlength{\parindent}{0em} % get rid of annoying indents
\setlength{\parskip}{0.5em}

\begin{document}
\maketitle

\begin{abstract}
\begin{quote}
Copyright \textcopyright  2018 CHIPFORGE.ORG. All rights reserved. \\

This process is licensed under the Libre Silicon public license; you can redistribute it and/or modify it under the terms of the Libre Silicon public license as published by the Libre Silicon alliance either version 1 of the License, or (at your option) any later version. \\

This design is distributed in the hope that it will be useful, but \textsc{WITHOUT ANY WARRANTY}; without even the implied warranty of \textsc{MERCHANTABILITY} or \textsc{FITNESS FOR A PARTICULAR PURPOSE}. See the Libre Silicon Public License for more details. \\

For further clarification consult the complete documentation of the process.
\end{quote}
\end{abstract}
\clearpage
\input{revision.tex}

PearlRiver is the first Test Wafer for the LibreSilicon 1 $ \mu m$ Technology.

This Testwafer contains almost all test structures to evaulate and qualify the technology-depended parameters.
On Wafer are placed structures to measure the sheet-resistance of all Poly and Metal Layer, as well Capacitors, NMOS and PMOS Transistors, Bipolar Junction Transistors, Diodes and Flash cells.
There are even common cells like NANDs, NORs and Inverters forming ring oscillators.

All parameters, measured from all test structures are used for Spice3f simulation models.

\clearpage
\tableofcontents
\clearpage

\pagestyle{headings}
\renewcommand{\theequation}{\thesection.\arabic{equation}}
\numberwithin{equation}{section}

\input{considerations.tex}
\input{ringoscillators.tex}

\input{teststructures.tex}

\end{document}
